---
title: "Elo linear analysis"
output: html_notebook
---

What's left to do:

- Pull latest project from the Github repository

- Create a folder called "combined_sets" in the "data" of the project

  - i.e. if you have the repository in a folder ~/ruslan/bu425-cluster/ then create a folder ~/ruslan/bu425-cluster/data/combined_sets/
  
- Download the datasets provided by Ruslan (https://drive.google.com/open?id=1CWd1yk89ibJcz4nhgj2Im97-crd_OmSw) - unzip the 6 datasets and put them in the combined_sets folder

- See notes below in second R block

- Add discussion to Google Drive 

  - Main points: Are there any statistically significant columns? Are there any highly correlated columns that we can remove to reduce complexity?

```{r}
library(leaps)
```

```{r}
data <- read.csv("../data/combined_sets/df_0.csv", header = TRUE)
for (idx in c(1,2,3,4,5)){
  data <- rbind(data, read.csv(sprintf("../data/combined_sets/df_%s.csv", idx), header = TRUE))
}

train <- read.csv("../data/train.csv", header = TRUE)

combined <- merge(x=data, y=train, by="card_id", all=TRUE)
# Filtered out any data that is NA
fdata <- combined[!is.na(combined$target),]

# Create categorial columns where appropriate
fdata$feature_1 <- as.factor(fdata$feature_1)
fdata$feature_2 <- as.factor(fdata$feature_2)
fdata$feature_3 <- as.factor(fdata$feature_3)regsubsets(Responsr_Var~. , data=" " , nbest=1)
```

```{r}
baseModel = lm(target ~ feature_1+feature_2+feature_3, data=fdata)
summary(baseModel)

allStateModel = lm(target ~ state_id_16+state_id_9+state_id_15+state_id_18+state_id_2+state_id_7+state_id_11+state_id_3+state_id_5+state_id_19+state_id_17+state_id_24+state_id_21+state_id_23+state_id_20+state_id_4+state_id_22+state_id_8+state_id_14+state_id_13+state_id_1+state_id_10+state_id_12+state_id_6+state_id_16+state_id_9+state_id_15+state_id_18+state_id_2+state_id_7+state_id_11+state_id_3+state_id_5+state_id_19+state_id_17+state_id_24+state_id_21+state_id_23+state_id_20+state_id_4+state_id_22+state_id_8+state_id_14+state_id_13+state_id_1+state_id_10+state_id_12+state_id_6, data=fdata)
summary(allStateModel)

allPurchaseModel = lm(target ~ avg_month_lag+avg_installments+avg_days_between_purch+std_days_between_purch+num_authorized+authorized_unauthorized_ratio+avg_purchase_amount+std_purchase_amount+min_purchase_amount+max_purchase_amount+num_purchases+feature_1+feature_2+feature_3, data=fdata)
summary(allPurchaseModel)
```